\documentclass[11pt,a4paper]{letter}
\usepackage[margin=1in]{geometry}
\usepackage{setspace}
\usepackage{hyperref}

\signature{Nguyen Nhat Huy\\International School, Duy Tan University\\Da Nang, Vietnam\\Email: nguyennhathuybdi06@gmail.com}

\date{\today}

\begin{document}

\begin{letter}{Editors of \textit{Discover Artificial Intelligence}\\Springer Nature}

\opening{Dear Editors of \textit{Discover Artificial Intelligence},}

We are pleased to submit our manuscript entitled:

\begin{center}
\textbf{``Insightimate: Enhancing Software Effort Estimation Accuracy Using Machine Learning Across Three Schemas (LOC/FP/UCP)''}
\end{center}

for consideration in \textit{Discover Artificial Intelligence}.

This manuscript presents a unified and reproducible machine-learning framework for software effort estimation, integrating three major sizing schemas (LOC, FP, UCP) within a single harmonized pipeline. Our study contributes a transparent cross-schema methodology, rigorous benchmarking across multiple ML models, and an enhanced COCOMO~II configuration. These contributions align with the journal’s focus on practical AI, interpretability, and reproducibility.

We believe our article is an excellent fit for \textit{Discover Artificial Intelligence} for the following reasons:

\begin{itemize}
    \item It provides a reproducible AI-based framework aligned with empirical best practices.
    \item It addresses a long-standing challenge in software engineering using modern AI techniques.
    \item It offers practical and industry-relevant insights for effort estimation and project prediction.
    \item All datasets used are publicly available and fully anonymized.
\end{itemize}

We further confirm that:

\begin{itemize}
    \item The manuscript has not been published before and is not under consideration elsewhere.
    \item All authors have read and approved the submitted version.
    \item There are no competing interests to declare.
    \item All data used in the study are publicly accessible.
\end{itemize}

Thank you very much for your consideration. We appreciate your time and look forward to the opportunity to revise and improve the manuscript as needed.

\closing{Sincerely,}

\end{letter}

\end{document}
